---
output:
  html_document:
    df_print: paged
---

STA360 Homework 1 (Ken Ye)

1.  Probability review:

    -   Exercise 2.2 from book

        a\.

        $$E[a_1Y_1+a_2Y_2] = E[a_1Y_1]+E[a_2Y_2] = a_1E[Y_1]+a_2E[Y_2] = a_1\mu_1+a_2\mu_2$$

        $$Var[a_1Y_1+a_2Y_2] = Var[a_1Y_1]+Var[a_2Y_2] = a_1^2Var[Y_1]+a_2^2Var[Y_2] = a_1^2\sigma_1^2+a_2^2\sigma_2^2$$

        b\.

        $$E[a_1Y_1-a_2Y_2] = E[a_1Y_1]-E[a_2Y_2] = a_1E[Y_1]-a_2E[Y_2] = a_1\mu_1-a_2\mu_2$$

        $$Var[a_1Y_1-a_2Y_2] = Var[a_1Y_1]+Var[a_2Y_2] = a_1^2Var[Y_1]+a_2^2Var[Y_2] = a_1^2\sigma_1^2+a_2^2\sigma_2^2$$

    -   Exercise 2.4 from book

        a\.

        $$
        Pr(H_j|E)Pr(E)= Pr(H_j\cap E) = Pr(E|H_j)Pr(H_j)
        $$ (by axiom P3)

        b\.

        $$
        Pr(E) = Pr(E\cap H_1) + Pr(E\cap H_2) + ... + Pr(E\cap H_k) = Pr(E\cap H_1) + Pr(E\cap \{\cup^{K}_{k=2}H_k\})
        $$

        (since $\{H_1,...,H_k\}$ is a partition of $H$)

        c\.

        $$
        Pr(E) = Pr(E\cap H_1) + Pr(E\cap H_2) + ... + Pr(E\cap H_k) = \sum^{K}_{k=1}Pr(E\cap H_k)
        $$

        (since $\{H_1,...,H_k\}$ is a partition of $H$)

        d\.

        $$
        \begin{aligned}
        Pr(H_j|E)Pr(E) &= Pr(E|H_j)Pr(H_j) \ \ \ \text{by a} \\
        Pr(H_j|E) &= \frac{Pr(E|H_j)Pr(H_j)}{Pr(E)} \\
        Pr(H_j|E) &= \frac{Pr(E|H_j)Pr(H_j)}{\sum^{K}_{k=1}Pr(E\cap H_k)} \ \ \ \text{by c}\\
        Pr(H_j|E) &= \frac{Pr(E|H_j)Pr(H_j)}{\sum^{K}_{k=1}Pr(E|H_k)Pr(H_k)} \\
        \end{aligned}
        $$

2.  Sensitivity and Specificity:

    a\. $$
    Pr(F) =  Pr(E)\times Pr(F|E) + Pr(F)\times Pr(F|not E) = 0.15\times0.94+0.85\times0.08 = 0.209 \\
    Pr(E|F) = \frac{Pr(F|E)Pr(E)}{Pr(F)} = \frac{0.94\times0.15}{0.209} \approx 0.67464
    $$

    b\.

    ```{r}
    interval <- seq(0, 1, length.out = 1000)

    # e is Pr(E)
    f <- function(e){
      return (0.94*e / (0.94*e + 0.08*(1-e)))
    }

    plot(interval, f(interval),
         main = 'Pr(E|F) VS Pr(E)',
         xlab = 'Pr(E)',
         ylab = 'Pr(E|F)')
    ```

3.  Joint distributions:

    a\.

    $h(y,\theta)$ is nonnegative because both $f(y,\theta)$ and $g(\theta)$ are nonnegative everywhere by defination as they are valid pdfs, implying their product $h(y,\theta)$ must be nonnegative everywhere as well.

    $$
    \int_\Theta \sum_{y\in \mathcal Y} h(y,\theta) d\theta = \int_\Theta \sum_{y\in \mathcal Y} f(y,\theta) g(\theta) d\theta = \int_\Theta 1\times g(\theta) d\theta = 1
    $$

    Therefore, $h(y,\theta)$ is a valid joint pdf on $\mathcal Y \times \Theta$.

    b\.

    $$
    p(y) = \int_0^\infty p(y,\theta) d \theta = \int_0^\infty f(y,\theta)g(\theta) d \theta \\
    p(\theta) = \sum_{y=0}^\infty p(y,\theta) = \sum_{y=0}^\infty f(y,\theta)g(\theta) = 1 \times g(\theta) = g(\theta)\\ 
    p(y|\theta) = \frac{p(y,\theta)}{p(\theta)} = \frac{f(y,\theta)g(\theta)}{\sum_{y=0}^\infty f(y,\theta)g(\theta)} = \frac{f(y,\theta)}{\sum_{y=0}^\infty f(y,\theta)} = \frac{f(y,\theta)}{1} = f(y,\theta)\\
    p(\theta|y) = \frac{p(y,\theta)}{p(y)} = \frac{f(y,\theta)g(\theta)}{\int_0^\infty f(y,\theta)g(\theta) d \theta}
    $$

4.  Posterior inference:

    a\.

    $$
    p(\theta|y = 2) = \frac{p(y = 2|\theta)p(\theta)}{p(y = 2)} = \frac{(5\theta)^2 e^{-5\theta}/2! \times \frac{1}{101}}{\sum_{\theta=0}^1 (5\theta)^2 e^{-5\theta}/2! \times \frac{1}{101}} = \frac{(5\theta)^2 e^{-5\theta}}{\sum_{\theta=0}^1 (5\theta)^2 e^{-5\theta}}
    $$

    ```{r}
    interval2 <- seq(0, 1, by = 1/100)

    f2 <- function(theta){
      nom <- ((5*theta)^2)*exp(-5*theta)
      denom <- sum(nom)
      return (nom/denom)
    }

    plot(interval2, f2(interval2),
         main = 'p(theta|y = 2) VS theta',
         xlab = 'theta',
         ylab = 'p(theta|y = 2)')
    ```

    b\.

    Under $p(\theta)$, median[$\theta$] = 50/100 = 1/2, mode[$\theta$] = {0, 1/100, 2/100, ... , 1}

    ```{r}
    vals <- f2(interval2)
    median_index <- which(cumsum(vals)>= 0.5)[1]
    median <- interval2[median_index]
    median
    ```

    ```{r}
    mode <- interval2[which.max(vals)]
    mode
    ```

    Under $p(\theta | y = 2)$, median[$\theta$] = 0.2, mode[$\theta$] = 0.4
